Philosophy of Distributional Semantics\\ \large Best Practices for the Application of Computational Methods to Questions of Meaning

### 2017-05-28 pds/Paper/paper_pds.tex
intro in decent shape

A third difficult point is the so-called \emph{double use} problem. New technology's use is often underspecified and could be used for benefit of humanity or to its detriment. Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media. \cite{•}
-Reeinforcing discrimination and bias

have the potential to amplify structural discrimination and systematically commit grave errors

Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases.
A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data.
Yet another issue is that machine learning techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization. \cite{barocas2016big}

\paragraph{2b moved}
Recently, machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.
%\begin{quote}
%In the years since, several disparate impact cases have made their way to the Supreme Court and lower courts, most having to do with employment discrimination. This June, the Supreme Court’s decision in Texas Dept. of Housing and Community Affairs v. Inclusive Communities Project, Inc. affirmed the use of the disparate impact theory to fight housing discrimination. The Inclusive Communities Project had used a statistical analysis of housing patterns to show that a tax credit program effectively segregated Texans by race. Sorelle Friedler, a computer science 
%researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
%\end{quote}

%\paragraph{Contribution}
%largest part of intro, preview of paper, methodology!
\paragraph{Public Attention is in ML minor extent in NLP, not in DS}
The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.

@@

We want to focus on aspects of discrimination and bias in the outcome because this seems least explored. Issues concerning privacy etc. seem to follow the same lines as general purpose ML, or even research, ethical principles while the issue of discrimination and bias has a special relation to DS.

\paragraph{Methodoology}

>>>>

A third difficult point is the so-called \emph{double use} problem. New technology's use is often underspecified and could be used for benefit of humanity or to its detriment. Chat-bots developed by researchers in automated natural language processing, to name a concrete example, can be used to assist people to access governmental health-care serices, yet they have reportedly been used by the Mexican government to disrupt political discussions deemed subversive on social media. \cite{leidner2017ethical}

Finally, machine learning harbours the potential to amplify structural discrimination and systematically disadvantage people because of their race, gender or other social category. If the models which are trained algorithmically are fed biased data, then predictions or classifications they make will reflect that bias.
Data processed by machine learning techniques often contains societal prejudice and bias, so that unquestioningly using them for decision making can severely reinforce society's existing biases. A related issue is that data available to the algorithms may systematically omit information relevant to what it is trying to uncover, as when minority groups are not adequately represented in the data. Yet another issue is that machine learning techniques may use useful patterns and relationships in the data that are really mirrorring historical factors of discrimination and marginalization. \cite{barocas2016big}

In the present essay, we will mostly be concerned with the last of these problematic issues surrounding machine learning.

%\begin{quote}
%In the years since, several disparate impact cases have made their way to the Supreme Court and lower courts, most having to do with employment discrimination. This June, the Supreme Court’s decision in Texas Dept. of Housing and Community Affairs v. Inclusive Communities Project, Inc. affirmed the use of the disparate impact theory to fight housing discrimination. The Inclusive Communities Project had used a statistical analysis of housing patterns to show that a tax credit program effectively segregated Texans by race. Sorelle Friedler, a computer science 
%researcher at Haverford College and a fellow at Data \& Society, called the Court’s decision “huge,” both “in favor of civil rights…and in favor of statistics.”
%\end{quote}

%\paragraph{Contribution}
%largest part of intro, preview of paper, methodology!
\paragraph{Public Attention is in ML minor extent in NLP, not in DS}
Many of the issues lately mentioned are of great importance to modern society at large. Consequently, with the rise of big data people working at the intersection of information technology and ethics have attempted to lay down foundations on how to handle these technologies. Many of these efforts have naturally been directed at the broad topic of machine learning in general.\cite{leidner2017ehtical} Only very recently the discussion has turned on machine learning used for analysis and modeling of language - often referred to as \emph{natural language processing}. Presumably because this is only one of many particular fields where machine learning is applied and the discussion of ethics natural language processing is bound pertain to less cases of applications throughout society, it took some time for the discussion to reach this field.

Recently, however, a discussion has opened up in a hitherto completely unmentioned domain. Machine learning techniques used for the computer-based study of language meaning - a field known as \emph{computational distributional semantics} - has come under scrutiny.
Various studies have shown that these methods yield results which are in important ways biased against particular groups as, for example, women \cite{google} \cite{wagner2015s}.The realization that algorithms fed on big data come with a host of problematic issues has led to the emergence of a variety of principles, concrete guidelines and public as well as institutional scrutiny.

@@

We want to focus on aspects of discrimination and bias in the outcome because these issues are least explored. Debates concerning privacy etc. would seem to follow the same lines as the discussion to establish general purpose ethical principles for machine learning, while the issue of discrimination and bias has a special relation to DS.

\paragraph{Methodoology}
The method employed by the present contribution is one of scanning previous efforts in a similar vein, identifying useful parts and applying them to ethical questions surrounding distributional semantics. There will be no foundational efforts undertaken to sustain the validity of the principles thus transferred.  Their validity is presupposed. The analysis contained in this contribution merely serves to disucss the usefulness of the principles in particular domain, and assert it where due.
